Forecasting the 2024 U.S. Presidential Election

### 2024-11-04 paper/paper.qmd

# Idealized Methodology

@@

```{r}
#| echo: false
#| warning: false
#| message: false
posterior_predictions <- posterior_predict(model, draws = 100)
ppc_dens_overlay(y = data_filtered$average_pct/100, yrep = posterior_predictions) +
  labs(title = "Posterior Predictive Check",
       x = "Rate of support",
       y = "Density")
```

@@

```{r}

>>>>

>>>>>>> cadfff05ead6aebff23ab03f96847c7f519121a6
# Idealized Methodology

@@

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows that the posterior predictive density closely follows the observed data density. Discrepancies between the two suggest that the model may be missing some aspect of the data structure or variability, which could guide potential model adjustments or refinement.

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows the prior for each parameter is centered around zero, but the posterior shows more precise estimates for each candidate and the date variable, reflecting the data's influence. Also, the posterior narrowing suggests that the model successfully extracted information from the data, reducing uncertainty around parameter estimates compared to the prior.
model_data <- read_parquet(here::here("data/02-analysis_data/model_data.parquet"))

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
pp_check(model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()

#| echo: false
#| warning: false
#| message: false
posterior_predictions <- posterior_predict(model, draws = 100)
ppc_dens_overlay(y = data_filtered$average_pct/100, yrep = posterior_predictions) +
  labs(title = "Posterior Predictive Check",
       x = "Rate of support",
       y = "Density")
```

@@

@fig-stanareyouokay-1 is a trace plot. It shows chain mixing, stationarity and no drifting. This suggests that the model's MCMC sampling achieved convergence for each parameter. The chains are well-mixed and stationary, providing confidence in the posterior estimates derived from this sampling process.

@fig-stanareyouokay-2 is a Rhat plot. It shows the Rhat values for the parameters estimated in the model. Values close to 1 indicate that the chains have mixed well and the posterior distribution has been adequately explored. This suggests that the model has converged successfully and the estimates are reliable for inference.

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(model, "trace")

plot(model, "rhat")
```
